---
title: Class Script
subtitle: "DANL 310: Data Visualization and Presentation"
author:
  - name: Byeong-Hak Choe
    affiliation: SUNY Geneseo
    
date: 2026-02-18

execute: 
  echo: true
  eval: true
  warning: false
  message: false
  fig-width: 10
  fig-height: 6


format:
  html:
    code-fold: true
    toc: true
    toc-title: "Contents"
    toc-expand: 3
    number-sections: true
    code-tools: true
    highlight-style: atom-one
---

```{r setup}
#| include: false

library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(DT)

# theme_set(theme_ipsum()+
#           theme(strip.background =element_rect(fill="lightgray"),
#                 axis.title.x =
#                   element_text(angle = 0,
#                                size = rel(1.33),
#                                margin = margin(10,0,0,0)),
#                 axis.title.y =
#                   element_text(angle = 90,
#                                size = rel(1.33),
#                                margin = margin(0,10,0,0))
#                 )
#           )
```


```{r}
library(tidyverse)
flights <- nycflights13::flights
```


# Classwork 6
## Q1
Find all flights that

```{r}
# Had an arrival delay of two or more hours
flights |> 
  filter(arr_delay > 120)

# Flew to Houston (IAH or HOU)
flights |> 
  filter(dest == "IAH" | dest == "HOU")

flights |> 
  filter(dest %in% c("IAH", "HOU"))

# Were operated by United (UA), American (AA), or Delta (DL)
flights |> 
  filter(carrier %in% c("UA", "AA", "DL"))

# Departed in summer (July, August, and September)
flights |> 
  filter(month %in% c(7,8,9))

# Arrived more than two hours late, but didn’t leave late
flights |> 
  filter(arr_delay > 120,
         dep_delay <= 0)

# Were delayed by at least an hour for the depature, but reduced the delay over 30 minutes during flight
flights |> 
  filter(dep_delay >= 60,
         dep_delay - arr_delay  > 30) |> 
  select(dep_delay, arr_delay)

# Departed between midnight and 6am (inclusive)

flights |> 
  filter( dep_time >= 0 , dep_time <= 600 ) |> 
  arrange(dep_time)


flights |> 
  filter( (dep_time >= 0 & dep_time <= 600) | dep_time == 2400 ) |> 
  arrange(-dep_time)

```


## Q2

How could you use `arrange()` to sort all missing values to the start?

```{r}
flights |> 
  skimr::skim(dep_delay)

flights |> 
  arrange(dep_delay)

```

```{r}

flights |> 
  arrange( desc( is.na(dep_delay) ) )

```

## Q3

Brainstorm as many ways as possible to select `dep_time`, `dep_delay`, `arr_time`, and `arr_delay` from flights.



```{r}
flights |> 
  select(starts_with("dep"), starts_with("arr"))
```


```{r}
flights |> 
  select(ends_with("_time"), ends_with("_delay"), -starts_with("sched"), -starts_with("air"))
```



## Q4

Find the 10 most delayed flights using a ranking function.

How do you want to handle ties?

```{r}
q4_arr <- flights |> 
  mutate(rank_min = min_rank(-arr_delay),
         rank_dense = dense_rank(-arr_delay),
         ) |> 
  arrange(rank_dense) |> 
  relocate(starts_with("rank"), arr_delay)
```

```{r}
q4_dep <- flights |> 
  mutate(rank_min = min_rank(-dep_delay),
         rank_dense = dense_rank(-dep_delay),
         ) |> 
  arrange(rank_dense) |> 
  relocate(starts_with("rank"), dep_delay)
```



## Q5
Which carrier has the worst arrival delays within each origin airport?

```{r}
q5 <- flights |> 
  group_by(origin) |> 
  filter(arr_delay == max(arr_delay, na.rm = T)) |> 
  relocate(origin, carrier, arr_delay)
  
```



## Q6

Which plane (`tailnum`) has the worst on-time arrival record?


```{r}
q6 <- flights |> 
  filter(arr_delay == max(arr_delay, na.rm = T)) |> 
  relocate(tailnum)
```



## Q7

What time of day should you fly if you want to avoid delays as much as possible?


```{r}

q7 <- flights |> 
  mutate(is_delayed = ifelse(arr_delay >= 30, T, F)) |> 
  count(hour, is_delayed) |> 
  group_by(hour) |> 
  mutate(prop = 100 * n / sum(n)) |> 
  filter(is_delayed) |> 
  arrange(prop)
```



## Q8

For each destination, compute the total minutes of delay. For each flight, compute the proportion of the total delay for its destination.


```{r}
q8 <- flights |> 
  group_by(dest) |> 
  mutate(tot_pos_dep_delay = sum(dep_delay[dep_delay > 0], na.rm = T),
         tot_pos_arr_delay = sum(arr_delay[arr_delay > 0], na.rm = T),
         tot_neg_dep_delay = sum(dep_delay[dep_delay < 0], na.rm = T),
         tot_neg_arr_delay = sum(arr_delay[arr_delay < 0], na.rm = T)
         ) |> 
  mutate(prop_pos_dep_delay = dep_delay / tot_pos_dep_delay,
         prop_pos_arr_delay = arr_delay / tot_pos_arr_delay,
         prop_neg_dep_delay = dep_delay / tot_neg_dep_delay,
         prop_neg_arr_delay = arr_delay / tot_neg_arr_delay,
         
         prop_pos_dep_delay = ifelse(prop_pos_dep_delay <= 0, NA, prop_pos_dep_delay),
         prop_pos_arr_delay = ifelse(prop_pos_arr_delay <= 0, NA, prop_pos_arr_delay),
         
         prop_neg_dep_delay = ifelse(prop_neg_dep_delay > 0, NA, prop_neg_dep_delay),
         prop_neg_arr_delay = ifelse(prop_pos_arr_delay > 0, NA, prop_neg_arr_delay)
         ) |> 
  select(dest, dep_delay, arr_delay, starts_with("tot"), starts_with("prop"))
```


## Q9

Find all destinations that are flown by at least two carriers. 

Use that information to rank the carriers.

```{r}
flights |> 
  count(dest) |> 
  nrow()
```


```{r}
q9 <- flights |> 
  count(dest, carrier) |> 
  group_by(dest) |> 
  filter( n() >= 2 ) |> 
  group_by(carrier) |> 
  summarize(tot_flights = sum(n),
            n_dest = n()) |> 
  arrange(-tot_flights)
```


## Q10
```{r}
tech_stocks <- read_csv(
  "https://bcdanl.github.io/data/tech_stocks_2015_2024.csv"
)
```
```{r}
tech_stocks |> 
  count(Ticker)
```
For each `Ticker`, compute the daily stock return as the percent change in the `Close` price from the previous trading day to the current day.

$$
(\text{Stock Return}_{d}) = \frac{(\text{Close}_{d}) - (\text{Close}_{d-1})}{(\text{Close}_{d-1})}
$$

```{r}
q10 <- tech_stocks |> 
  group_by(Ticker) |> 
  mutate(Close_lag = lag(Close)) |> 
  select(Date, Ticker, starts_with("Close"), Volume) |> 
  mutate(return = 100 * (Close - Close_lag) / Close_lag,
         return_rounded = round(return, 2))
```



## Q11

```{r}
q10 |> 
  ggplot(aes(x = return, y = Volume)) +
  geom_point(alpha = .1) +
  geom_smooth()
```
```{r}
q10 |> 
  ggplot(aes(x = return, y = Volume,
             color = Ticker,
             fill = Ticker)) +
  geom_point(alpha = .1) +
  geom_smooth()
```

# Scale

```{r}
library(gapminder)

gapminder <- gapminder

```




```{r}
gapminder |>
  filter(year == 2007) |>
  ggplot(aes(x = gdpPercap, 
             y = lifeExp,
             color = continent)) +
  geom_point(alpha = 0.8,
             size = 3) +
  geom_smooth(se = F) +
  scale_x_continuous(
    breaks = c(1000, 10000, 50000),
    labels = scales::dollar
  ) +
  labs(
    x = "GDP per capita",
    y = "Life expectancy",
    color = "Continent")
```


