---
title: Class Script
subtitle: "DANL 310: Data Visualization and Presentation"
author:
  - name: Byeong-Hak Choe
    affiliation: SUNY Geneseo
    
date: 2026-03-09

execute: 
  echo: true
  eval: true
  warning: false
  message: false
  fig-width: 10
  fig-height: 6


format:
  html:
    code-fold: true
    toc: true
    toc-title: "Contents"
    toc-expand: 3
    number-sections: true
    code-tools: true
    highlight-style: atom-one
---

```{r setup}
#| include: false

library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(DT)

# theme_set(theme_ipsum()+
#           theme(strip.background =element_rect(fill="lightgray"),
#                 axis.title.x =
#                   element_text(angle = 0,
#                                size = rel(1.33),
#                                margin = margin(10,0,0,0)),
#                 axis.title.y =
#                   element_text(angle = 90,
#                                size = rel(1.33),
#                                margin = margin(0,10,0,0))
#                 )
#           )
```

```{r}
library(tidyverse)
gss_cat <- gss_cat
```


# fct_reorder2
```{r}
by_age <- gss_cat |>
  filter(!is.na(age)) |>
  count(age, marital) |>
  group_by(age) |>
  mutate(prop = n / sum(n))

# Default legend order (alphabetical)
ggplot(by_age, 
       aes(x = age, y = prop, 
           color = marital)) +
  geom_line(linewidth = 1) +
  scale_color_brewer(palette = "Set1")

```


```{r}
#| fig-height: 10
# Legend order follows the line endings at the largest age
ggplot(by_age, aes(x = age, y = prop,
                   color = fct_reorder2(marital, age, prop))) +
  geom_line(linewidth = 1) +
  scale_color_brewer(palette = "Set1") +
  labs(color = "marital")
```

# fct_relevel
```{r}
rincome_summary <- gss_cat |>
  group_by(rincome) |>
  summarize(
    age = mean(age, na.rm = T),
    tvhours = mean(tvhours, na.rm = T),
    n = n()
  )
```


```{r}
ggplot(rincome_summary, 
       aes(age, fct_reorder(rincome, age) ) )  + 
  geom_point()
```
```{r}
levels(rincome_summary$rincome)
```


```{r}
ggplot(rincome_summary, 
       aes(age, 
           fct_relevel(rincome, 
                       "Not applicable") ) )  +  # making it to the first level
  geom_point()
```


# str_c()

```{r}
str_length(c("a", "R for data science", NA))

str_c("x", "y", "z")

str_c(c("x", "y", "z"), 
      sep = ", ")


str_c("prefix-", c("a", "b", "c"), "-suffix")


str_c(c("x", "y", "z"), collapse = ", ")
```


# str_split()

```{r}
sentences
```

```{r}
sentences |>
  head(5) |> 
  str_split(" ")
```
```{r}
x <- c("Apple", "Banana", "Pear")
str_sub(x, 1, 3)

## negative numbers count backwards from end
str_sub(x, -3, -1)
```


```{r}
x <- c("apple", "banana", "pear")

str_detect(x, "e")

str_replace(x, "a", "-")

str_replace_all(x, "a", "-")
```



# Classwork 8

```{r}
uk_soccer <- read_csv("https://bcdanl.github.io/data/premier_league_2022.csv")

rmarkdown::paged_table(uk_soccer)
```


## Q1a

```{r}
tott_home <- uk_soccer |> 
  filter(HomeTeam == "Tottenham") |> 
  select(FTR, HTR)

tott_away <- uk_soccer |> 
  filter(AwayTeam == "Tottenham") |> 
  select(FTR, HTR)
```


## Q1b
Create the following four data.frames.
`home_htr` that counts the number of observations for each value of HTR in tott_home.
`home_ftr` that counts the number of observations for each value of FTR in tott_home.
`away_htr` that counts the number of observations for each value of HTR in tott_away.
`away_ftr` that counts the number of observations for each value of FTR in tott_away.
```{r}
home_htr <- tott_home |> 
  count(HTR)

home_ftr <- tott_home |> 
  count(FTR)

away_htr <- tott_away |> 
  count(HTR)

away_ftr <- tott_away |> 
  count(FTR)

```


## Q1c

Create the following two data.frames:
`home_results` is created using `home_ftr` and `home_htr`;
`away_results` is created using `away_ftr` and `away_htr`.

```{r}
home_results <- cbind(home_ftr, 
                      home_htr)

colnames(home_results) <- c("result", "FTR", "tmp", "HTR")

home_results <- home_results|> 
  mutate(tott_location = "Home",
         .before = 0) |> 
  select(-tmp) 
```


```{r}
away_results <- cbind(away_ftr, 
                      away_htr)

colnames(away_results) <- c("result", "FTR", "tmp", "HTR")

away_results <- away_results|> 
  mutate(tott_location = "Away",
         .before = 0) |> 
  select(-tmp) 
```

## Q1d
```{r}
home_results <- home_results |> 
  mutate(result = case_when(
    result == "A" ~ "Lose",
    result == "D" ~ "Draw",
    result == "H" ~ "Win"
  ))
```


```{r}
away_results <- away_results |> 
  mutate(result = case_when(
    result == "A" ~ "Win",
    result == "D" ~ "Draw",
    result == "H" ~ "Lose"
  ))
```


## Q1e

```{r}
home_results_longer <- home_results |> 
  pivot_longer(cols = FTR:HTR,
               names_to = "time",
               values_to = "count")
```

```{r}
away_results_longer <- away_results |> 
  pivot_longer(cols = FTR:HTR,
               names_to = "time",
               values_to = "count")
```

```{r}
tott_results <- rbind(home_results_longer,
                      away_results_longer)
```


## Q1f

```{r}
tott_results |> 
  mutate(time = ifelse(time == "FTR", 
                       "Full-Time", "Half-Time"),
         time = factor(time,
                       levels = c("Full-Time",
                                  "Half-Time")
                       ),
         result = factor(result,
                         levels = c("Win", "Draw", "Lose"))
         ) |> 
  ggplot(aes(x = result,
             y = count,
             fill = time)) +
  geom_col(position = "dodge") +
  facet_wrap(~tott_location) +
  scale_fill_colorblind() +
  theme_minimal() +
  theme(
    legend.position = "top",
    legend.title.position = "top",
    legend.title = element_text(hjust = .5,
                                face = "bold"),
    legend.text = element_text(hjust = .5,
                                face = "italic"),
    strip.background = element_rect(fill = "hotpink",
                                    color = "black")
  ) +
  guides(
    fill = guide_legend(
      label.position = "bottom",
      keywidth = 3
    )
  )
```



