---
title: "March 31, 2025 - Class Script"
subtitle: "DANL 310: Introduction to Data Analytics"
author: 
  - name: Byeong-Hak Choe
date: 2025-03-31

execute: 
  echo: true
  eval: true
  warning: false
  message: false
  fig-width: 9
  # fig-height: 5
  
format: 
  html:
    toc: true
    number-sections: true
    code-fold: show       # https://quarto.org/docs/output-formats/html-code.html  
    code-tools: true      # https://quarto.org/docs/reference/cells/cells-jupyter.html
    highlight-style: atom-one # atom-one tango espresso
---

```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(DT)
library(socviz) # textbook r package
# To set default ggplot theme
# theme_set(theme_fivethirtyeight() +
#           theme(strip.background = element_rect(fill="lightgray"),
#                 axis.title.x = 
#                   element_text(angle = 0,
#                                size = rel(1.33),
#                                margin = margin(10,0,0,0)),
#                 axis.title.y = 
#                   element_text(angle = 90,
#                                size = rel(1.33),
#                                margin = margin(0,10,0,0))
#                 )
          # ) # theme_minimal() becomes default
# scale_color_discrete <- function(...) scale_color_tableau(...)
# scale_fill_discrete <- function(...) scale_fill_tableau(...)
```

# Classwork 7 - Question 4

```{r}
holiday_movies <- read_csv("https://bcdanl.github.io/data/holiday_movies.csv")
holiday_movie_genres <- read_csv("https://bcdanl.github.io/data/holiday_movie_genres.csv")
```

## Q4a

Provide the R code to generate the data.frame, holiday_movie_with_genres, which combines the two data.frames, holiday_movies and holiday_movie_genres:


```{r}
holiday_movie_with_genres <- holiday_movie_genres |> 
  left_join(holiday_movies)
```






## Q4b
Provide the R code using skimr::skim() to see how the summary statistics—mean, median, standard deviation, minimum, maximum, first and third quartiles—of `average_rating` and `num_votes` varies by popular genres and `title_type`.

Consider only the five popular genres, which are selected in terms of the number of titles for each `genre`.
Removes the video type of the titles when calculating the summary statistics.

```{r}
popular_genres <- holiday_movie_genres |> 
  count(genres) |> 
  mutate(ranking = dense_rank(-n)) |> 
  filter(ranking <= 5)

unique(holiday_movie_with_genres$title_type)


q4b <- holiday_movie_with_genres |> 
  filter(genres %in% popular_genres$genres,
         title_type != "video") |> 
  group_by(genres, title_type) |> 
  skim(average_rating, num_votes)

```



## Q4c

Provide R code to recreate the ggplot figure illustrating how the relationship between log10(num_votes) and average_rating varies by the popular genres and title_type.

The five popular genres are selected in terms of the number of titles for each genre.
The video type of the titles are removed in the ggplot figure.


```{r}
holiday_movie_with_genres |> 
  filter(genres %in% popular_genres$genres,
         title_type != "video") |> 
  ggplot(aes(x = log10(num_votes),
             y = average_rating)) +
  geom_point(alpha = .2) +
  geom_smooth(aes(color = genres,
                  fill = genres),
              method = 'lm') +
  facet_grid(title_type ~ genres,
             scale = 'free_y')
```


## Q4d
Provide a comment to illustrate how the relationship between log10(num_votes) and average_rating varies by the popular genres and title_type.

- For Romance, the association is positive, for both movie and tvMovie
- For Family, the association is positive for movie, but is negative for tvMovie.







