---
title: "Untitled"
format: html
editor: visual
---


# Visualizing New York City with **ggmap** and OpenStreetMap data  
This walkthrough shows how to use **ggmap** (leveraging free OpenStreetMap tiles) to  

1. display a street map of Manhattan,  
2. mark **Central Park** and **The Metropolitan Museum of Art**, and  
3. draw a **driving route** from **Wall Street** to **Times Square**.

> **Packages used:** `ggmap`, `ggplot2`, `dplyr`, `sf`, and `osrm`  
> – `ggmap` fetches the base map  
> – `osrm` (Open Source Routing Machine) returns an OSM‑based route without the need for a Google API key  
> – `sf` handles the route geometry

---

## 1. Setup

```{r}
# install.packages(c("ggmap", "osrm", "sf", "dplyr"))
library(ggmap)
library(dplyr)
library(sf)
library(osrm)
```



```{r}
# Rough center of Manhattan
nyc_center <- c(lon = -74.00, lat = 40.72)

# A zoom level of 12–13 shows all requested locations clearly
nyc_map <- get_map(location = nyc_center,
                   zoom      = 12,
                   source    = "osm")  # fetches OpenStreetMap tiles
```

